---
title: "lab_02"
author: "Judith Altneu"
date: "2024-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1. 26 departments offer courses with climate in the name. Atmospheric and Oceanic Science offered the most classes with climate in the title with 68.**

```{r}

 courses_climate <- umd_courses |>
  filter(str_detect(title, "Climate")) |> filter(seats >= 1) |>
 group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))

```

**Q2.** Of the departments listed in A2, which one is the most interesting to you, and why? **A2. I thought Criminology and Criminal Justice was interesting since they offer two courses related to climate. I have taken several criminology classes and none have really connected to climate related themes.**

**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result?

**A3. Business and Management offered the most courses in Fall 2023 with 164. Business and Management remained at the top with 82 classes having at least one syllabus posted. School of Music and Information Science classes remained in the top five but Economics and Biological Sciences Program had a lot of classes with syllabi posted.**

```{r}
 umd_courses |>
  filter(term == 202308) |>
 group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))

 umd_courses |>
  filter(term == 202308 & syllabus_count >= 1) |>
 group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))
  
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time. **A4. Wes Moore's campaign increased spending each month and spent the most money in the months leading up to the election particularly the summer and October.**

```{r}
 moore_campaign_expenses_by_month <- maryland_expenses |> 
  mutate(expenditure_date_new =mdy(expenditure_date)) |>
    mutate(month = floor_date(expenditure_date_new, "month")) |>
  filter(committee_name == 'Moore  Wes For Maryland') |>
  group_by(month) |>
  summarize(total_amount = sum(amount))

```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5 About \$25 million more was spent in Maryland compared to donations coming from out of state.**

```{r}
maryland_expenses_with_state <- maryland_expenses |>
  mutate(
    spending_location = case_when(
        str_detect(address, " Maryland ") ~ "In-state",
         .default = "Out of state"
      )
  )
maryland_expenses_with_state |>
    group_by(spending_location) |>
  summarize (total_amount = sum(amount, na.rm=T)) 
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6.** For the Maryland expenditure data how many people contributed over \$1000?

I found that there are over 18,000 entries of people that donated over \$1000. I guess I could have made the amount I was filtering by a lot bigger but still didn't think a ton of people donated over \$1000.

```{r}
contriubtions_over_1000 <- maryland_expenses |>
   filter(amount >= 1000) |>
    arrange(desc(amount))
  
```
